---
title: "Technology, Quality of Life, and Relationships"
output: pdf_document
---

```{r loading libraries, echo=FALSE}
#Loading libraries
library(ggplot2)
library(gridExtra)
library(scales)
library(plyr)

#Loading Data
df <- read.csv("May_2013_Online_Dating_CSV.csv")
```

```{r gender, echo=FALSE}
# Gender plot

okdf <- df

okdf[, "sex"] <- as.factor(okdf[, "sex"])
#summary(okdf[, "sex"])
#str(okdf[, "sex"])

gender <- ggplot(data = okdf, aes(x = sex, fill = sex)) + 
  geom_bar() + scale_x_discrete(labels = c("1" = "Male", "2" = "Female")) +
  labs(x = "Gender", y = "Count", title = "Gender")
```

```{r sexual orientation, echo=FALSE}
# Sexual orientation plot
okdf[, "lgbt"] <- as.factor(okdf[, "lgbt"])
#summary(okdf$lgbt)
#str(okdf$lgbt)

sexual_orientation <- ggplot(data = okdf, aes(x = lgbt)) + 
  geom_bar(fill='lightgreen',color='black') + 
  scale_x_discrete(labels = c("1" = "Heterosexual", "2" = "Gay/lesbian", "3" = "Bisexual", "4" = "Other", "8" = "Don't know", "9" = "Refused")) + coord_flip() +
  labs(x = "", y = "Count", title = "Sexual Orientation")
  
```

```{r Hispanic, echo=FALSE}
# Hispanic
okdf[, "hisp"] <- as.factor(okdf[, "hisp"])
#summary(okdf$hisp)
#str(okdf$hisp)

hispanic <- ggplot(data = okdf, aes(x = hisp)) + 
  geom_bar(fill='lightgreen',color='black') + 
  scale_x_discrete(labels = c("1"="Yes", "2"="No", "8" = "Don't know", "9" ="Refused")) +
  labs(x = "", y = "Count", title = "Latino Background")
```

```{r age, echo=FALSE}
#Age plot
#summary(okdf$age)
#str(okdf$age)
median.age <- median(df$age)

age <- ggplot(data = okdf, aes(x = age)) + 
  geom_histogram(binwidth = 5, color="black", fill="orange") + 
  scale_x_continuous(limits = c(15,100), breaks = seq(0, 100, 10)) +
  labs(x = "Age", y = "Frequency", title = "Age")

```


```{r race, echo=FALSE}
# Race plot
okdf[, "race"] <- as.factor(okdf[, "race"])
#summary(okdf$race)
#str(okdf$race)

r_race <- ggplot(data = okdf, aes(x=reorder(race, -table(race)[race]))) + geom_bar(fill = "lightgreen", color = "black") + 
  scale_x_discrete(labels=c("1" = 'White',"2" = 'Black',"3" = 'Asian',"4" = 'Mixed',"5" = 'Native American',"6" = 'Other',"8" = "Don't Know","9" = 'Refused')) +
  labs(x = "", y = "Count", title = "Race") +
  coord_flip()
```


```{r employment, echo=FALSE}
# Employment plot
okdf[, "emplnw"] <- as.factor(okdf[, "emplnw"])
#summary(okdf$emplnw)
#str(okdf$emplnw)
retired.percent <- 616/2252
retired.percent <- 942/2252


employment <- ggplot(data = okdf, aes(x=reorder(emplnw, -table(emplnw)[emplnw]))) + 
  geom_bar(fill = "lightgreen", colour = "black") + 
  scale_x_discrete(labels=c("1" = 'Full-time',"2" = 'Part-time',"3" = 'Retired',"4" = 'Not employed',"5" = 'Self-employed',"6" = 'Disabled',"7" = "Student", "8" = 'Other', "9" = "Refused")) +
  labs(x = "", y = "Count", title = "Employment Level") + coord_flip()
```


```{r income, echo=FALSE}
# Income plot
okdf[, "inc"] <- as.factor(okdf[, "inc"])
#summary(okdf$inc)
#str(okdf$inc)

income.refused <- 299/2252

income <- ggplot(data = okdf, aes(x = inc)) + 
  geom_bar(fill = "lightgreen", colour = "black") + coord_flip() +
  scale_x_discrete(labels = c("1" = "under 10k", "2" = "10-20k", "3" = "20-30k", "4" = "30-40k", "5" = "40-50k", "6" = "50-75k", "7" = "75-100k", "8" = "100-150k", "9" = "over 150k", "98" = "Don't Know", "99" = "Refused")) +
  labs(x = "", y = "Count", title = "Income in thousands USD")
  
```

```{r education, echo=FALSE}
# Education plot
okdf[, "educ2"] <- as.factor(okdf[, "educ2"])
#summary(okdf$educ2)
#str(okdf$educ2)

education <- ggplot(data = okdf, aes(x = educ2)) + 
  geom_bar(fill = "lightgreen", colour = "black") + 
  scale_x_discrete(labels = c("1" = "Grade 1-8", "2" = "Incomplete high school", "3" = "Complete high school", "4" = "Some college", "5" = "Associate degree", "6" = "Bachelor's", "7" = "Some postgrad", "8" = "Master's/phd/professional", "98" = "Don't know", "99" = "Refused")) +
  labs(x = "", y = "Count", title = "Education") + coord_flip()
```


```{r household size, echo=FALSE}
# Household size plot
okdf[, "hh1"] <- as.factor(okdf[, "hh1"])
#summary(okdf$hh1)
#str(okdf$hh1)

household_size <- ggplot(data = okdf, aes(x = hh1)) + 
  geom_bar(fill = "lightgreen", colour = "black") + 
  scale_x_discrete(labels = c("1" = "1", "2" = "2", "3" = "3", "4" = "4", "5" = "5", "6" = "6+", "8" = "Don't \n know", "9" = "Refu-\n sed")) +
  labs(x = "", y = "Number of Respondents", title = "Household Size")
```


```{r marrital status, echo=FALSE}
# Marital status
okdf[, "mar"] <- as.factor(okdf[, "mar"])
#summary(okdf$mar)
#str(okdf$mar)
married.percent <- 1093/2552

marital_status <- ggplot(data = okdf, aes(x = mar)) + 
  geom_bar(fill = "lightgreen", colour = "black") + 
  scale_x_discrete(labels = c("1"="Married", "2"="Living with Partner", "3"="Divorced", "4"= "Separated", "5" = "Widowed", "6" = "Never been married", "8" = "Don't know", "9" ="Refused")) +
  labs(x = "", y = "Count", title = "Marital status") + coord_flip()
```


```{r children, echo=FALSE}
# Children
okdf[, "par"] <- as.factor(okdf[, "par"])
#summary(okdf$par)
#str(okdf$par)

children <- ggplot(data = okdf, aes(x = par)) + 
  geom_bar(fill = "lightgreen", colour = "black") + 
  scale_x_discrete(labels = c("1"="Yes", "2"="No", "8" = "Don't know", "9" ="Refused")) +
  labs(x = "", y = "Number of Respondents", title = "Children <18 living in household")
```


```{r life quality, echo=FALSE}
# Life quality
okdf[, "q1"] <- as.factor(okdf[, "q1"])
#summary(okdf$q1)
#str(okdf$q1)

life_quality <- ggplot(data = okdf, aes(x = q1)) + geom_bar(fill = "lightgreen", colour = "black") +
  scale_x_discrete(labels = c("1"="Excellent", "2"="Very good", "3" = "Good", "4" = "Fair", "5" = "Poor", "8" = "Don't know", "9" ="Refused")) +
  labs(x = "", y = "Count", title = "Quality of life") + coord_flip()
```


###  **Introduction**   

Our dataset is compiled of categorical data received through telephone interviews in the USA in 2013. The dataset was originally used for a study on online dating; however through investigation of the variables, insight can be generated on the quality of life for a sample of Americans.

With our core research topic focusing on the sample’s quality of life, the data offers two sections of analysis: firstly through relationships and secondly through Home Economics. 

The following slides will provide an overview of the dataset.
  
  
#### **Decriptive Statistics (part 1)**


As seen from below, the respondents were predominantly white, slightly older (median age of 53).

```{r describtive_stats_1, echo=FALSE , warning=FALSE}

grid.arrange(gender, r_race, age, hispanic)
```


#### **Decriptive Statistics (part 2)**


The respondents are well educated, with 42% employed full time; however, the number of retirees is also large (27% of all respondents). The majority live in a 2-person household. It is noteworthy to point out that 299 people (13%) refused to disclose their family income level. The second most heard response (also 13%) was earnings of 50-75 thousand USD per year.

```{r describtive_stats_2, echo=FALSE , warning=FALSE}
grid.arrange(employment, education, income, household_size)
```


#### **Decriptive Statistics (part 3)**


The respondents are predominantly heterosexual; 43% are married. A greater amount of respondents do not have dependants under 18 living in their homes. The quality of life of the respondents resembles a normal distribution.

```{r describtive_stats_3, echo=FALSE, warning=FALSE}
grid.arrange(marital_status, sexual_orientation, children, life_quality)
```





### **Relationships & Quality of Life**

Within our project, we observed that online dating is a focus of many of the questions. As the objective of our analysis is one of a modern perception of quality of life, and online dating is such a novel phenomenon, we believe it fitting to explore how relationships affect quality of life with a certain focus on online dating.

On the graphs below, we explored the relationship between:

* The search for a romantic partner and having an online dating profile, and
* The perception of how difficult it is to meet a partner and an online dating profile

Limitation 1: while there were a total of 138 people in our sample looking for a partner, only 51 answered the question on whether they have an online dating profile. The situation was similar for people that weren't looking for a partner. Thus, the following graphs only contain responses for people that answered both questions.

Limitation 2: while there were a total of 134 people in our sample who answered the question on how  easy/difficult it is to meet people, only 50 answered the question on whether they have an online dating profile. 

* According to the left-most graph, of 51 people looking for a romantic partner, 33  had an active online dating profile, whereas, 18 individuals did not. 

* On the right-most graph, of 50 respondents, 58% felt it was difficult to meet people in the city or town where they lived. Of this subsample, 62% had a currently posted online profile. This is however a  small and specific subsection of the greater interviewed sample, so other related questions with greater response rates will be investigated.

```{r Relationships & Quality of Life, echo=FALSE, warning=FALSE}

# 3. For people who say that they are looking for a partner, look at the proportion of who are using online dating

lookmeet <- df$mar4
lookmeet <-  ifelse(lookmeet>2,NA,lookmeet)
usedod = df$date6
usedod = ifelse(usedod>2,NA,usedod)
dtf = as.data.frame(table(lookmeet,usedod))

online.profile.looking <- ggplot(dtf,aes(x=lookmeet,y=Freq)) + geom_bar(stat="identity",aes(fill=usedod)) + scale_x_discrete(labels = c("1"="Looking for Patner", "2"="Not looking for Partner")) + scale_fill_discrete(name="Current Online Dating Profile",labels=c("Yes","No")) + labs(x = "", y = "Number of Respndents", title = "Search for Romantic Partner\n & Online Dating") + theme(legend.position="bottom")

# 5 For people that say it’s difficult / easy to meet people in their city/state, see how many people using online dating?

meet = df$meet
meet = ifelse(meet>2,NA,meet)
prof = df$date6
prof = ifelse(prof>2,NA,prof)
dtf3 = as.data.frame(table(meet,prof))

online.profile.meeting <- ggplot(dtf3,aes(x=meet,y=Freq)) + geom_bar(stat="identity",aes(fill=prof)) + scale_x_discrete(labels = c("1"="Easy to meet", "2"="Difficult to meet")) + scale_fill_discrete(name="Current Online Dating Profile",labels=c("Yes","No")) + labs(x="",y="Number of Respondents", title = "Perceived Ease of Meeting People\n & Online Dating") + theme(legend.position="bottom")

grid.arrange(online.profile.looking, online.profile.meeting, ncol = 2)

```


* of the individuals that find it difficult to date 27% currently had an online dating profile

* Whereas, for individuals that find it easy to date in their city, 20% were actively using an online profile.

```{r echo=FALSE, warning=FALSE}

c1 = table(df$celldatea,df$celldateb)
c1vec = as.vector(c1[1:2,1:2])
names(c1vec) = c("calling / texting", "calling","texting","neither")
c1df = data.frame(c1vec)
p1 = ggplot(c1df,aes(x = rownames(c1df),y = c1vec)) + geom_bar(stat="identity",fill='lightblue',color='black') + labs(x="Using Phones to meet people",y="Number of people")+geom_text(aes(label=c1vec), vjust=1.2, color='darkblue')

search = df$intdatec
search = ifelse(search>2,NA,search)
askout = df$intdated
askout = ifelse(askout>2,NA,askout)
dtf1 = as.data.frame(table(search,askout))
p2 = ggplot(dtf1,aes(x=search,y=Freq))+geom_bar(stat="identity",aes(fill=askout))+scale_x_discrete(labels = c("1"="Search online", "2"="Don't search online"))+scale_fill_discrete(name="Ask out",labels=c("Yes","No"))+labs(x="Using Internet to meet people",y="Number of People")

follow = df$snsdate1e
follow = ifelse(follow>2,NA,follow)
ask = df$snsdate1c
ask = ifelse(ask>2,NA,ask)
dtf2 = as.data.frame(table(follow,ask))
p3 = ggplot(dtf2,aes(x=follow,y=Freq))+geom_bar(stat="identity",aes(fill=ask))+scale_x_discrete(labels = c("1"="Follow/friend", "2"="Don't follow/friend"))+scale_fill_discrete(name="Ask out",labels=c("Yes","No"))+labs(x="Using Social Media to meet people ",y="Number of People")

g1 = table(df$date5a,df$date5b)
g1vec = as.vector(g1[1:2,1:2])
names(g1vec) = c("used OD / got married","used OD","got married","neither")
g1df = data.frame(g1vec)
p4 = ggplot(g1df,aes(x = rownames(g1df),y = g1vec)) + geom_bar(stat="identity",fill='lightpink',color='black') + labs(x="Using Online Dating (OD) to meet people",y="Number of people")+geom_text(aes(label=g1vec), vjust=1.2, color='black')
 
grid.arrange(p1,p2,p3,p4,nrow=2)
```


Four ideas – what proportion of people that have used phone, internet, social media or online dating to meet people.

* People prefer internet to meet people as opposed to phones. Most individuals in the sample use the internet to search for prospective people to meet, whereas on the phone respondents prefer to ask people out directly by calling them.
 *	Of the people that used the internet as means of dating (found in the lower two graphs)
 *Social media in absolute terms is a reasonably unpopular means of dating with only 27 people using social media to friend/follow people and ask them out on dates.
 *	The greatest use of social media as dating technology tool, is to simply follow or friend individuals without asking them out on a date.
 *	On the bottom right graph it can be seen that 33% of people that met met people through an online dating website got married or entered a long-term relationship.


#### **Origin of the Relationship vs Quality of Life**


Based on the graph below, it may seem that the origine of the relationship (whether a person met their partner online or offline) has some association to people's quality of life. For example: 

* 35% of people that met offline said their quality of life was good. Meanwhile, only 27% of people that met online believed the same. 
* 20% of people that met offline said their quality of life was excellent. Meanwhile, 27% of people that met online believed the same.

However, running the chi squre test does not support the claim that the two distributions are different (p value is high).

```{r echo = FALSE, warning = FALSE} 


indhap <- which(df$q1 < 6) 
indmet <- which(df$rel1 == 1 |df$rel1 == 2)

finind <- intersect(indhap,indmet)

hap <- df[finind, "q1"]
met <- df[finind, "rel1"]

ex<- cbind.data.frame(hap,met)

ex1df <- data.frame(ex)

ex1df <- as.data.frame(with(ex, table(hap,met)))

ex1df[,"met"] <- revalue(ex1df[,"met"], c("1"="Met Online", "2" = "Met offline"))

ex1df <- ddply(ex1df, .(met), transform, p = Freq/sum(Freq))

# happiness vs how people met
ggplot(data = ex1df , aes(x = hap, y = p)) + geom_bar(stat = "identity", fill = "lightgreen", color = "black") + facet_grid(~ met) + scale_x_discrete(labels = c("1"="Excellent", "2"="Very good", "3" = "Good", "4" = "Fair", "5" = "Poor")) + labs( x = "Quality of Life" , y = "Percentage", title = "Quality of life vs Origin of Relationship") + geom_text(aes(label = Freq), vjust=-.5)

chisq.test(ex[,"met"],ex[,"hap"])

```


#### **Privacy & Conflict in a Relationship**


We can use logistic regression to find the impact of having a common email account, a shared social networking profile, a shared online calendar, and shared online passwords on conflicts in relationships. 

```{r echo = FALSE , warning= FALSE} 

#8.  Relationship between privacy [REL4 and REL5] and having a fight in a relationship [REL8]? Does person who shares password/account together tend to have more fights than those who are not?

library(caTools)
df1 = subset(df, rel6 == 1 | rel6 == 2)
df1$rel6 = ifelse(df1$rel6==1,0,1) #convert rel6 into a binomial variable
set.seed(1000)
split = sample.split(df1$rel6,SplitRatio=0.7) # divide the data into training and test sets, with 70% data in the training set
train = subset(df1,split==TRUE) #Training set
test = subset(df1, split==FALSE)# Test set
model1 = glm(rel6~rel4a+rel4b+rel4c + rel5,data=train,family=binomial)# Logistic Regression model
summary(model1)
```

Currently 115 people say that internet and email has had a major impact on their relationship, while 213 people say that it has had a negative impact. So, for our baseline model, we will assume that internet always has a negative impact, and out baseline accuracy will be 115/115+213  = 35%. Can we come up with a logistic regression model which improves this accuracy number?


### **Quality of Life and Household Economics**


The second aspect of this study focuses on how household economics, specifically – family structure, education, income and available technology – relates to the respondent’s quality of life


#### **Quality of Life vs presence of children in the household**


Looking at the graph below, it appears that households without children are less likely to say that their quality of life is Excellent. However, using the chi square test, we see that the p-value is high, indicating that there is not enough evidence to conclude that there are differences in quality of life among people with and without children <18 in the household. 

Note: This dataset exludes 32 respondents who answered "Don't Know" or "Refuse" to either the question on the quality of life or the question on whether they have children under 18 in the household.

```{r life quality vs children, echo=FALSE}
# Life quality vs children

# data frame for children and quality of life

quality_vs_children <- df[df$par!="8" & df$par!="9" & df$q1!="9" & df$q1!="8",]

quality_vs_children_plot <- as.data.frame(with(quality_vs_children, table(par,q1)))

quality_vs_children_plot <- ddply(quality_vs_children_plot, .(par), transform, p = Freq/sum(Freq))

quality_vs_children_plot[,"q1"] <- revalue(quality_vs_children_plot[,"q1"], c("1"= "Excellent", "2" = "Very Good", "3" = "Good", "4" = "Fair", "5" = "Poor"))

quality_vs_children_plot[,"par"] <- revalue(quality_vs_children_plot[,"par"], c("1" = "Children <18 = Yes", "2" = "Children <18 = No"))

ggplot(quality_vs_children_plot, aes(q1, p)) + geom_bar(stat = "identity", fill = "lightgreen", color = "black") + facet_grid(~par) + labs(x = "Quality of Life", y = "Percent of Respondents", title = "Quality of Life vs Presence of Children under 18 in the Household") + geom_text(aes(label = Freq), vjust=-.5)

chisq.test(quality_vs_children[,"par"], quality_vs_children[,"q1"])
```


#### **Quality of Life vs Use of Cell Phones**


Looking at the graph below, it appears that respondents that use cellphones are more likely to say that their quality of life is Excellent. Using the chi square test, we see that the p-value is low, indicating that there is evidence to conclude that there are differences in the quality of life levels among people that use internet on the phone and those that don't. 

This dataset exludes 22 respondents who answered "Don't Know" or "Refuse" to either the question on the quality of life or the question on whether they use internet on their cellphone.


```{r life quality vs internet on cell phones, echo=FALSE}
# Life quality vs use of internet on cellphone
# summary(df[,"intmob"])

quality_vs_phoneNet <- df[df$intmob!="8" & df$q1!="9" & df$q1!="8",]

quality_vs_phoneNet_plot <- as.data.frame(with(quality_vs_phoneNet, table(intmob,q1)))

quality_vs_phoneNet_plot <- ddply(quality_vs_phoneNet_plot, .(intmob), transform, p = Freq/sum(Freq))

quality_vs_phoneNet_plot[,"q1"] <- revalue(quality_vs_phoneNet_plot[,"q1"], c("1"= "Excellent", "2" = "Very Good", "3" = "Good", "4" = "Fair", "5" = "Poor"))

quality_vs_phoneNet_plot[,"intmob"] <- revalue(quality_vs_phoneNet_plot[,"intmob"], c("1"="Internet on Cellphone = Yes", "2" = "Internet on Cellphone = No"))
  
ggplot(quality_vs_phoneNet_plot, aes(q1, p)) + geom_bar(stat = "identity", fill = "lightgreen", color = "black") + facet_grid(~intmob) + labs(x = "Quality of Life", y = "Percent of Respondents", title = "Quality of Life vs Use of Internet on Cellphone") + geom_text(aes(label = Freq), vjust=-.5)

chisq.test(quality_vs_phoneNet[,"intmob"], quality_vs_phoneNet[,"q1"])
```


#### **Quality of Life vs Use of Social Networks**


Looking at the graph below, the users and non-users of social networks (Facebook, Twitter, LinkedIn, etc) appear to report their quality of life levels similarly. Using the chi square test, we see that the p-value is high, indicating that there is no evidence to conclude that there are differences in the quality of life levels between users and non-users of social networks. 

Note: This dataset exludes 371 respondents who answered "Don't Know" or "Refuse" to either the question on the quality of life or the question on the use of social networks. Respondents who do not use internet were also exlcluded.

```{r life quality vs social networks, echo=FALSE}
# Life quality vs social networks
# summary(df[,"act87"])

# summary(df[,"act112"])

quality_vs_socNetworks <- df[df$act87!="9" & !is.na(df$act87) & df$act112!="9" & !is.na(df$act112) & df$act112!="8" & df$q1!="9" & df$q1!="8",]

quality_vs_socNetworks <- subset(quality_vs_socNetworks, select = c(act87, act112, q1))

for (i in 1:nrow(quality_vs_socNetworks)){
  if (quality_vs_socNetworks[i,"act87"]=="1" |  quality_vs_socNetworks[i,"act112"]=="1"){
    quality_vs_socNetworks[i,"SN_user"] <- "Social Network User = Yes"
  }
  else{
    quality_vs_socNetworks[i,"SN_user"] <- "Social Network User = No"
  }
}

quality_vs_socNetworks_plot <- as.data.frame(with(quality_vs_socNetworks, table(SN_user,q1)))

quality_vs_socNetworks_plot <- ddply(quality_vs_socNetworks_plot, .(SN_user), transform, p = Freq/sum(Freq))

quality_vs_socNetworks_plot[,"q1"] <- revalue(quality_vs_socNetworks_plot[,"q1"], c("1"= "Excellent", "2" = "Very Good", "3" = "Good", "4" = "Fair", "5" = "Poor"))

ggplot(quality_vs_socNetworks_plot, aes(q1, p)) + geom_bar(stat = "identity", fill = "lightgreen", color = "black") + facet_grid(~SN_user) + labs(x = "Quality of Life", y = "Percent of Respondents", title = "Quality of Life vs Use of Social Networks")

chisq.test(quality_vs_socNetworks[,"SN_user"], quality_vs_socNetworks[,"q1"])
```


#### **Income vs Smart Phone Use**


The following graph demonstrates the relationship between type of smartphone used and income. Visually, it appears that the prevalence of the iPhone is associated with higher income level. 


```{r income vs smart phones, echo=FALSE, warning = FALSE}
# Group discrete variables for smartphone (smart2) and income (inc)

shdf <- read.csv("May_2013_Online_Dating_CSV.csv")

shdf[,"smart2"] <- as.numeric(shdf[,"smart2"])
shdf [!is.na(shdf$smart2) & shdf$smart2 > 5,]$smart2 <- 9
shdf [!is.na(shdf$inc) & shdf$inc < 4,]$inc <- 1
shdf [!is.na(shdf$inc) & shdf$inc > 3 & shdf$inc < 7,]$inc <- 2
shdf [!is.na(shdf$inc) & shdf$inc > 6 & shdf$inc < 9,]$inc <- 3
shdf [!is.na(shdf$inc) & shdf$inc > 10,]$inc <- 99
shdf[,"smart2"] <- as.character(shdf[,"smart2"])
shdf[,"inc"] <- as.character(shdf[,"inc"])

ggplot(data = subset(shdf, !is.na(shdf$smart2)),aes (x = inc, fill = smart2, order = -as.numeric(smart2))) + geom_bar () + scale_x_discrete(labels = c("1"="<30K", "2"="30K - 75K", "3"="75K - 150K", "9"=">150K", "99" = "N/A")) + scale_fill_discrete(name = "Smartphone", breaks = c ("1", "2", "3", "4", "9"), labels = c("iPhone", "Blackberry", "Android", "Windows", "Others")) + labs(x = "Income (in US$)", y = "Number of Respondents", title = "Smartphone & Income") 

```

#### **Income vs Education**

Looking at the graph below, it appears that there may be a difference among income levels across different education levels. Chi square test and low p-value confirm this.

```{r income vs education, echo=FALSE}
#Kirby's data frame
kcdf <- df

# use smartphone/Internet vs. edu level
use.smartphone <- as.factor(kcdf[ , 'q12'])
use.internet <- as.factor(kcdf[ , 'intuse'])
edu <- as.factor(kcdf[ , 'educ2'])
kcdf.edu <- as.data.frame(cbind(use.internet,use.smartphone))
kcdf.edu <- cbind(kcdf.edu,edu)


# edu vs. income 
income <- as.factor(kcdf[ , 'inc'])
kcdf.income <- cbind(kcdf.edu,income)

# remove all 98 and 98, which stand for don't know and don't want to tell
kcdf.income <- subset(kcdf.income, kcdf.income[ ,'edu'] != 98)
kcdf.income <- subset(kcdf.income, kcdf.income[ ,'edu'] != 99)
kcdf.income <- subset(kcdf.income, kcdf.income[ ,'income']!= 98)
kcdf.income <- subset(kcdf.income, kcdf.income[ ,'income']!= 99)

income.edu <- kcdf.income[ , c('edu','income')]

# setting level names for income
income.edu[,"income"] <- revalue(income.edu[,"income"], c("1" = "under 10k", "2" = "10-20k", "3" = "20-30k", "4" = "30-40k", "5" = "40-50k", "6" = "50-75k", "7" = "75-100k", "8" = "100-150k", "9" = "over 150k"))

ggplot(income.edu, aes(edu, fill=income) ) +
  geom_bar(position="fill") +
  scale_x_discrete(labels = c("1" = "Grade 1-8", "2" = "Incomplete high school", "3" = "Complete high school", "4" = "Some college", "5" = "Associate degree", "6" = "Bachelor's", "7" = "Some postgrad", "8" = "Master's/phd/professional")) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(x = "", y = "Number of Respondents", title = "Income vs Education")

```


### **Limitations**

While the population size was composed of 2252 respondents, due to significant clustering of respondents into groups (internet users, mobile internet users, social network users, etc), the number of responds upon whom analysis of online dating behaviour could be completed was very small. 

* There are only  198 people in the sample that used online dating sites 
* Only 57 people have used a dating app on their phone

As a result, limited amout of research could be done specifically on online dating behavior and how it affects quality of life. 

To study this phenomena at greater depth, a recommendaiton for the future analysis would be to collect data solely from online dating service users. 

The fact that the dataset contained no continuous variables, limited the type of analysis and modelling that could be completed. 


### **Conclusion**

To be written.